### 2017-10-24 Draft2/Appending.Dynamic.tex
Remove a cdot
Language
Remove all the \varnitems stuff

			\State $\ListName.\FieldHead[\ListName.\FieldHeadSize] \gets item$
			\State $\ListName.\FieldSize \gets \ListName.\FieldSize + 1$
		\EndProcedure
		\Statex
		\Procedure{$\FuncGrow$}{$L$}
			\If{$\ListName.\FieldEmpty$}
				\State $\ListName.\FieldBuffer \gets \FuncNewArray(\InitCapacity)$
			\Else
				\State $new\ buf \gets \FuncNewArray(\ListName.\FieldSize \cdot \GrowthFactor)$
				\State $\FuncArrayCopy(source: \ListName.\FieldBuffer, dest: new\ buf, len: \ListName.\FieldSize)$
				\State $\ListName.\FieldBuffer \gets new\ buf$
			\EndIf
		\EndProcedure
	\end{algorithmic}
\end{algorithm}

@@

Before I analyze the time complexity of $\FuncAppend$, I consider a different method for measuring its cost. Suppose I start with an empty list and $\varnitems$ elements are appended. How many times is an element stored in an array? I will term the answer to this question the \textbf{write cost} of $\varnitems$ appends, and denote it $\nwritesfn(n)$.

In the above code, one write is performed after the conditional, so it is apparent that $\nwritesfn(\varnitems) \geq \varnitems$ after $n$ appends. However, $\FuncGrow$ also does some writing, so in order to find a precise formula for $\nwritesfn(\varnitems)$, I need to analyze when $\FuncGrow$ is called. Consider the following lemma:

\begin{lemma}
\label{CapacitySequence}
	Let $\ListName$ by a dynamic array. As $\varnitems$ items are appended, the sequence of capacities $\ListName$ takes on is $$S_C = 0,\ \InitCapacity,\ \GrowthFactor\InitCapacity,\ \GrowthFactor^2\InitCapacity,\ \ldots\ \GrowthFactor^{\lceil \log_{\GrowthFactor} \varnitems - \log_{\GrowthFactor} \InitCapacity \rceil}\InitCapacity$$.
\end{lemma}

\begin{proof}
	The capacity starts out at zero and becomes $\InitCapacity$ once one element is appended. From then on, it can only grow by a factor of $\GrowthFactor$ via $\FuncGrow$, so if $\GrowthFactor^i\InitCapacity$ is the current capacity then $\GrowthFactor^{i + 1}\InitCapacity$ must be the next capacity. By induction, the rest of the sequence is $$\left\{ \GrowthFactor^i\InitCapacity \right\}_{i = 1}^k$$ for some $k$. Since the capacity is only as big as is needs to be, namely greater than or equal to $\varnitems$, $k$ is the smallest integer for which $\GrowthFactor^k\InitCapacity \geq \varnitems$. Using this property, I now derive $k$.

	\begin{align*}
	\GrowthFactor^k\InitCapacity &\geq \varnitems\\
	\GrowthFactor^k &\geq \frac{\varnitems}{\InitCapacity}\\
	k &\geq \log_{\GrowthFactor} \varnitems - \log_{\GrowthFactor} \InitCapacity\\
	k &\geq \log_{\GrowthFactor} \varnitems - \log_{\GrowthFactor} \InitCapacity > k - 1\\
	k &= \big\lceil \log_{\GrowthFactor} \varnitems - \log_{\GrowthFactor} \InitCapacity \big\rceil
	\end{align*}

	Then the final term in the sequence is $\GrowthFactor^{\lceil \log_{\GrowthFactor} \varnitems - \log_{\GrowthFactor} \InitCapacity \rceil}\InitCapacity$, completing the proof.
\end{proof}

[make lemma]
Now, consider that to grow by $\GrowthFactor$, $\FuncGrow$ must be called. So $\GrowthFactor^k\InitCapacity$ being present in $S_C$ means $\FuncGrow$ was called at size $\GrowthFactor^{k - 1}\InitCapacity$ for $k \geq 1$. Then $\FuncGrow$ was called at the sizes $$S_R = 0,\ \InitCapacity,\ \GrowthFactor\InitCapacity,\ \GrowthFactor^2\InitCapacity,\ \ldots\ \GrowthFactor^{\lceil \log_{\GrowthFactor} \varnitems - \log_{\GrowthFactor} \InitCapacity \rceil - 1}\InitCapacity$$

Each time $\FuncGrow$ is called, $i$ items are copied where $i$ is the current size of the dynamic array. Then the total number of items copied for $\varnitems$ appends is

\begin{align*}
0 + \InitCapacity + \GrowthFactor\InitCapacity + \ldots + \GrowthFactor^{\lceil \log_{\GrowthFactor} \varnitems - \log_{\GrowthFactor} \InitCapacity \rceil - 1}\InitCapacity &= \left( \frac{\GrowthFactor^{\lceil \log_{\GrowthFactor} \varnitems - \log_{\GrowthFactor} \InitCapacity \rceil} - 1}{\GrowthFactor - 1} \right) \InitCapacity
\end{align*}

Finally, appending the writes made for every item in $\FuncAppend$, I finally find an explicit formula for $\nwritesfn(\varnitems)$.

[ref1]
$$
\nwritesfn(\varnitems) = \varnitems + \left( \frac{\GrowthFactor^{\lceil \log_{\GrowthFactor} \varnitems - \log_{\GrowthFactor} \InitCapacity \rceil} - 1}{\GrowthFactor - 1} \right) \InitCapacity
$$

[ref2]
To make this expression easier to work with, I will write it in big-P notation to approximate it for large $\varnitems$. First I note that $\log_{\GrowthFactor} k \leq \lceil \log_{\GrowthFactor} k \rceil < \left( \log_{\GrowthFactor} k \right) + 1$ for any $k > 0$, so $k \leq \GrowthFactor^{\lceil \log_{\GrowthFactor} k \rceil} < \GrowthFactor{k}$. Then

\begin{align*}
\nwritesfn(\varnitems) &= \varnitems + \left( \frac{\GrowthFactor^{\big\lceil \log_{\GrowthFactor} \left( \frac {\varnitems} {\InitCapacity} \right) \big\rceil} - 1}{\GrowthFactor - 1} \right) \InitCapacity\\
\varnitems + \left( \frac{\frac{\varnitems}{\InitCapacity} - 1}{\GrowthFactor - 1} \right) \InitCapacity \leq \nwritesfn(\varnitems) &< \varnitems + \left( \frac{\GrowthFactor\frac{\varnitems}{\InitCapacity} - 1}{\GrowthFactor - 1} \right) \InitCapacity\\
\varnitems + \left( \frac{\varnitems - \InitCapacity}{\GrowthFactor - 1} \right) \leq \nwritesfn(\varnitems) &< \varnitems + \left( \frac{\GrowthFactor\varnitems - \InitCapacity}{\GrowthFactor - 1} \right)\\
P\left( \varnitems + \left( \frac{\varnitems - \InitCapacity}{\GrowthFactor - 1} \right) \right) \leq P\left( \nwritesfn(\varnitems) \right) &< P\left( \varnitems + \left( \frac{\GrowthFactor\varnitems - \InitCapacity}{\GrowthFactor - 1} \right) \right)\\
P\left( \left( \frac{\GrowthFactor}{\GrowthFactor - 1} \right) \varnitems \right) \leq P\left( \nwritesfn(\varnitems) \right) &< P\left( \left( \frac{2\GrowthFactor - 1}{\GrowthFactor - 1} \right) \varnitems \right)
\end{align*}

@@

I wish to find the space allocated while appending $\varnitems$ items to a dynamic array. I denote this quantity $\spacefn(\varnitems)$, and define it as the total length of all arrays allocated during $\varnitems$ $\FuncAppend$ calls. I now derive a formula for $\spacefn(\varnitems)$.

From \ref{CapacitySequence} I know the exact capacities $\ListName$ takes on as $\varnitems$ items are appended. Dynamic arrays also have the property that $i \in S_C$ iff an array of length $i$ was allocated. Then the total length of arrays allocated is

\begin{align*}
\spacefn(\varnitems) &= \sum_{i \in S_C} i\\
&= 0 + \InitCapacity + \GrowthFactor\InitCapacity + \GrowthFactor^2\InitCapacity + \ldots + \GrowthFactor^{\lceil \log_{\GrowthFactor} \varnitems - \log_{\GrowthFactor} \InitCapacity \rceil}\InitCapacity\\
&= \left( \frac{\GrowthFactor^{\lceil \log_{\GrowthFactor} \varnitems - \log_{\GrowthFactor} \InitCapacity \rceil + 1} - 1}{\GrowthFactor - 1} \right) \InitCapacity
\end{align*}

@@

\begin{align*}
\left( \frac{\GrowthFactor \frac{\varnitems}{\InitCapacity} - 1}{\GrowthFactor - 1} \right) \InitCapacity \leq \spacefn(\varnitems) &< \left( \frac{\GrowthFactor^2 \frac{\varnitems}{\InitCapacity} - 1}{\GrowthFactor - 1} \right) \InitCapacity\\
P\left( \left( \frac{\GrowthFactor \frac{\varnitems}{\InitCapacity} - 1}{\GrowthFactor - 1} \right) \InitCapacity \right) \leq P\left( \spacefn(\varnitems) \right) &< P\left( \left( \frac{\GrowthFactor^2 \frac{\varnitems}{\InitCapacity} - 1}{\GrowthFactor - 1} \right) \InitCapacity \right)\\
P\left( \left( \frac{\GrowthFactor}{\GrowthFactor - 1} \right) \varnitems \right) \leq P\left( \spacefn(\varnitems) \right) &< P\left( \left( \frac{\GrowthFactor^2}{\GrowthFactor - 1} \right) \varnitems \right)
\end{align*}

>>>>

			\State $\ListName.\FieldHead[\ListName.\FieldHeadSize] \gets item$
			\State $\ListName.\FieldSize \gets \ListName.\FieldSize + 1$
		\EndProcedure
		\Statex
		\Procedure{$\FuncGrow$}{$L$}
			\If{$\ListName.\FieldEmpty$}
				\State $\ListName.\FieldBuffer \gets \FuncNewArray(\InitCapacity)$
			\Else
				\State $new\ buf \gets \FuncNewArray(\ListName.\FieldSize \times \GrowthFactor)$
				\State $\FuncArrayCopy(source: \ListName.\FieldBuffer, dest: new\ buf, len: \ListName.\FieldSize)$
				\State $\ListName.\FieldBuffer \gets new\ buf$
			\EndIf
		\EndProcedure
	\end{algorithmic}
\end{algorithm}

@@

Before I analyze the time complexity of $\FuncAppend$, I consider a different method for measuring its cost. Suppose I start with an empty list and $n$ elements are appended. How many times is an element stored in an array? I will term the answer to this question the \textbf{write cost} of $n$ appends, and denote it $\nwritesfn(n)$.

In the above code, one write is performed after the conditional, so it is apparent that $\nwritesfn(n) \geq n$ after $n$ appends. However, $\FuncGrow$ also does some writing, so in order to find a precise formula for $\nwritesfn(n)$, I need to analyze when $\FuncGrow$ is called. Consider the following lemma:

\begin{lemma}
\label{CapacitySequence}
	Let $\ListName$ by a dynamic array. As $n$ items are appended, the sequence of capacities $\ListName$ takes on is $$S_C = 0,\ \InitCapacity,\ \GrowthFactor\InitCapacity,\ \GrowthFactor^2\InitCapacity,\ \ldots\ \GrowthFactor^{\lceil \log_{\GrowthFactor} n - \log_{\GrowthFactor} \InitCapacity \rceil}\InitCapacity$$.
\end{lemma}

\begin{proof}
	The capacity starts out at zero and becomes $\InitCapacity$ once one element is appended. From then on, it can only grow by a factor of $\GrowthFactor$ via $\FuncGrow$, so if $\GrowthFactor^i\InitCapacity$ is the current capacity then $\GrowthFactor^{i + 1}\InitCapacity$ must be the next capacity. By induction, the rest of the sequence is $$\left\{ \GrowthFactor^i\InitCapacity \right\}_{i = 1}^k$$ for some $k$. Since the capacity is only as big as is needs to be, namely greater than or equal to $n$, $k$ is the smallest integer for which $\GrowthFactor^k\InitCapacity \geq n$. Using this property, I now derive $k$.

	\begin{align*}
	\GrowthFactor^k\InitCapacity &\geq n\\
	\GrowthFactor^k &\geq \frac{n}{\InitCapacity}\\
	k &\geq \log_{\GrowthFactor} n - \log_{\GrowthFactor} \InitCapacity\\
	k &\geq \log_{\GrowthFactor} n - \log_{\GrowthFactor} \InitCapacity > k - 1\\
	k &= \big\lceil \log_{\GrowthFactor} n - \log_{\GrowthFactor} \InitCapacity \big\rceil
	\end{align*}

	Then the final term in the sequence is $\GrowthFactor^{\lceil \log_{\GrowthFactor} n - \log_{\GrowthFactor} \InitCapacity \rceil}\InitCapacity$, completing the proof.
\end{proof}

[make lemma]
Now, consider that to grow by $\GrowthFactor$, $\FuncGrow$ must be called. So $\GrowthFactor^k\InitCapacity$ being present in $S_C$ means $\FuncGrow$ was called at size $\GrowthFactor^{k - 1}\InitCapacity$ for $k \geq 1$. Then $\FuncGrow$ was called at the sizes $$S_R = 0,\ \InitCapacity,\ \GrowthFactor\InitCapacity,\ \GrowthFactor^2\InitCapacity,\ \ldots\ \GrowthFactor^{\lceil \log_{\GrowthFactor} n - \log_{\GrowthFactor} \InitCapacity \rceil - 1}\InitCapacity$$

Each time $\FuncGrow$ is called, $i$ items are copied where $i$ is the current size of the dynamic array. Then the total number of items copied for $n$ appends is

\begin{align*}
0 + \InitCapacity + \GrowthFactor\InitCapacity + \ldots + \GrowthFactor^{\lceil \log_{\GrowthFactor} n - \log_{\GrowthFactor} \InitCapacity \rceil - 1}\InitCapacity &= \left( \frac{\GrowthFactor^{\lceil \log_{\GrowthFactor} n - \log_{\GrowthFactor} \InitCapacity \rceil} - 1}{\GrowthFactor - 1} \right) \InitCapacity
\end{align*}

Finally, counting the writes made for every item in $\FuncAppend$, I finally find an explicit formula for $\nwritesfn(n)$.

[ref1]
$$
\nwritesfn(n) = n + \left( \frac{\GrowthFactor^{\lceil \log_{\GrowthFactor} n - \log_{\GrowthFactor} \InitCapacity \rceil} - 1}{\GrowthFactor - 1} \right) \InitCapacity
$$

[ref2]
To make this expression easier to work with, I will write it in big-P notation to approximate it for large $n$. First I note that $\log_{\GrowthFactor} k \leq \lceil \log_{\GrowthFactor} k \rceil < \left( \log_{\GrowthFactor} k \right) + 1$ for any $k > 0$, so $k \leq \GrowthFactor^{\lceil \log_{\GrowthFactor} k \rceil} < \GrowthFactor{k}$. Then

\begin{align*}
\nwritesfn(n) &= n + \left( \frac{\GrowthFactor^{\big\lceil \log_{\GrowthFactor} \left( \frac {n} {\InitCapacity} \right) \big\rceil} - 1}{\GrowthFactor - 1} \right) \InitCapacity\\
n + \left( \frac{\frac{n}{\InitCapacity} - 1}{\GrowthFactor - 1} \right) \InitCapacity \leq \nwritesfn(n) &< n + \left( \frac{\GrowthFactor\frac{n}{\InitCapacity} - 1}{\GrowthFactor - 1} \right) \InitCapacity\\
n + \left( \frac{n - \InitCapacity}{\GrowthFactor - 1} \right) \leq \nwritesfn(n) &< n + \left( \frac{\GrowthFactorn - \InitCapacity}{\GrowthFactor - 1} \right)\\
P\left( n + \left( \frac{n - \InitCapacity}{\GrowthFactor - 1} \right) \right) \leq P\left( \nwritesfn(n) \right) &< P\left( n + \left( \frac{\GrowthFactorn - \InitCapacity}{\GrowthFactor - 1} \right) \right)\\
P\left( \left( \frac{\GrowthFactor}{\GrowthFactor - 1} \right) n \right) \leq P\left( \nwritesfn(n) \right) &< P\left( \left( \frac{2\GrowthFactor - 1}{\GrowthFactor - 1} \right) n \right)
\end{align*}

@@

I wish to find the space allocated while appending $n$ items to a dynamic array. I denote this quantity $\spacefn(n)$, and define it as the total length of all arrays allocated during $n$ $\FuncAppend$ calls. I now derive a formula for $\spacefn(n)$.

From \ref{CapacitySequence} I know the exact capacities $\ListName$ takes on as $n$ items are appended. Dynamic arrays also have the property that $i \in S_C$ iff an array of length $i$ was allocated. Then the total length of arrays allocated is

\begin{align*}
\spacefn(n) &= \sum_{i \in S_C} i\\
&= 0 + \InitCapacity + \GrowthFactor\InitCapacity + \GrowthFactor^2\InitCapacity + \ldots + \GrowthFactor^{\lceil \log_{\GrowthFactor} n - \log_{\GrowthFactor} \InitCapacity \rceil}\InitCapacity\\
&= \left( \frac{\GrowthFactor^{\lceil \log_{\GrowthFactor} n - \log_{\GrowthFactor} \InitCapacity \rceil + 1} - 1}{\GrowthFactor - 1} \right) \InitCapacity
\end{align*}

@@

\begin{align*}
\left( \frac{\GrowthFactor \frac{n}{\InitCapacity} - 1}{\GrowthFactor - 1} \right) \InitCapacity \leq \spacefn(n) &< \left( \frac{\GrowthFactor^2 \frac{n}{\InitCapacity} - 1}{\GrowthFactor - 1} \right) \InitCapacity\\
P\left( \left( \frac{\GrowthFactor \frac{n}{\InitCapacity} - 1}{\GrowthFactor - 1} \right) \InitCapacity \right) \leq P\left( \spacefn(n) \right) &< P\left( \left( \frac{\GrowthFactor^2 \frac{n}{\InitCapacity} - 1}{\GrowthFactor - 1} \right) \InitCapacity \right)\\
P\left( \left( \frac{\GrowthFactor}{\GrowthFactor - 1} \right) n \right) \leq P\left( \spacefn(n) \right) &< P\left( \left( \frac{\GrowthFactor^2}{\GrowthFactor - 1} \right) n \right)
\end{align*}

### 2017-10-24 Draft2/Appending.Growth.tex
Remove all the \varnitems stuff

I again start off by finding the write cost when $\varnitems$ are appended. I give the write cost function a slightly different name, $\nwritesnewfn(\varnitems)$, so I can compare it to $\nwritesfn(\varnitems)$ later. I also use $\GrowthFactornew$ and $\InitCapacitynew$ instead of $\GrowthFactor$ and $\InitCapacity$ for the growth factor and initial capacity, respectively.

@@

$$
S_R = 0,\ \InitCapacitynew,\ \GrowthFactornew\InitCapacitynew,\ \GrowthFactornew^2\InitCapacitynew,\ \ldots\ \GrowthFactornew^{\lceil \log_{\GrowthFactornew} \varnitems - \log_{\GrowthFactornew} \InitCapacitynew \rceil - 1}\InitCapacitynew
$$

Let $k$ be the number of times the tail is appended to. Since the tail is a dynamic array, the number of writes it makes is $\nwritesfn(k)$. Since $\FuncGrow$ is called $|S_R|$ times and it appends to the tail each time except for the first, $k = |S_R| - 1 = \lceil \log_{\GrowthFactornew} \varnitems - \log_{\GrowthFactornew} \InitCapacitynew \rceil$. Then the formula for $\nwritesgrowfn(n)$ is

@@

\begin{align*}
k &= \lceil \log_{\GrowthFactornew} \varnitems - \log_{\GrowthFactornew} \InitCapacitynew \rceil\\
O(k) &= O(\log \varnitems)
\end{align*}

Then, it is shown that the $k$-term disappears when analyzing the big-P complexity of $\nwritesnewfn(\varnitems)$, leaving just $P(\varnitems)$:

\begin{align*}
P(\nwritesnewfn(\varnitems)) &= P(\varnitems + \nwritesfn(k))\\
&= P(\varnitems) + P(\nwritesfn(k))\\
&= P(\varnitems) + O(\log \varnitems)\\
&= P(\varnitems)
\end{align*}

@@

% Should this be C(n) instead of C?
The space function for $\varnitems$ $\FuncAppend$ calls on a growth array is denoted $\spacenewfn(\varnitems)$. Since growth arrays never throw away item buffers, if the current capacity is $C$ then the total length of item buffers is also $C$. $\spacenewfn(\varnitems)$ is usually greater than $C$, however; this is because growth arrays not only allocate item buffers, but stores references to them in the tail. Thus the space the tail allocates must also be determined.

I determined in [...] that if $\varnitems$ are appended then the tail is appended to $k$ times, where $k = \lceil \log_{\GrowthFactornew} \varnitems - \log_{\GrowthFactornew} \InitCapacitynew \rceil$. Since the tail is a dynamic array, it allocates $\spacefn(k)$ space, where $\spacefn$ is the space function of the tail. Through a similar method to [...] it can be shown $O(\spacefn(k)) = O(k) = O(\log \varnitems)$. Then % TODO: P-analysis should come after the explicit formula.

\begin{align*}
\spacenewfn(\varnitems) = P(2^{\lceil \log_2 \varnitems \rceil}) = O(\varnitems)
\end{align*}

>>>>

I again start off by finding the write cost when $n$ are appended. I give the write cost function a slightly different name, $\nwritesnewfn(n)$, so I can compare it to $\nwritesfn(n)$ later. I also use $\GrowthFactornew$ and $\InitCapacitynew$ instead of $\GrowthFactor$ and $\InitCapacity$ for the growth factor and initial capacity, respectively.

@@

$$
S_R = 0,\ \InitCapacitynew,\ \GrowthFactornew\InitCapacitynew,\ \GrowthFactornew^2\InitCapacitynew,\ \ldots\ \GrowthFactornew^{\lceil \log_{\GrowthFactornew} n - \log_{\GrowthFactornew} \InitCapacitynew \rceil - 1}\InitCapacitynew
$$

Let $k$ be the number of times the tail is appended to. Since the tail is a dynamic array, the number of writes it makes is $\nwritesfn(k)$. Since $\FuncGrow$ is called $|S_R|$ times and it appends to the tail each time except for the first, $k = |S_R| - 1 = \lceil \log_{\GrowthFactornew} n - \log_{\GrowthFactornew} \InitCapacitynew \rceil$. Then the formula for $\nwritesgrowfn(n)$ is

@@

\begin{align*}
k &= \lceil \log_{\GrowthFactornew} n - \log_{\GrowthFactornew} \InitCapacitynew \rceil\\
O(k) &= O(\log n)
\end{align*}

Then, it is shown that the $k$-term disappears when analyzing the big-P complexity of $\nwritesnewfn(n)$, leaving just $P(n)$:

\begin{align*}
P(\nwritesnewfn(n)) &= P(n + \nwritesfn(k))\\
&= P(n) + P(\nwritesfn(k))\\
&= P(n) + O(\log n)\\
&= P(n)
\end{align*}

@@

% Should this be C(n) instead of C?
The space function for $n$ $\FuncAppend$ calls on a growth array is denoted $\spacenewfn(n)$. Since growth arrays never throw away item buffers, if the current capacity is $C$ then the total length of item buffers is also $C$. $\spacenewfn(n)$ is usually greater than $C$, however; this is because growth arrays not only allocate item buffers, but stores references to them in the tail. Thus the space the tail allocates must also be determined.

I determined in [...] that if $n$ are appended then the tail is appended to $k$ times, where $k = \lceil \log_{\GrowthFactornew} n - \log_{\GrowthFactornew} \InitCapacitynew \rceil$. Since the tail is a dynamic array, it allocates $\spacefn(k)$ space, where $\spacefn$ is the space function of the tail. Through a similar method to [...] it can be shown $O(\spacefn(k)) = O(k) = O(\log n)$. Then % TODO: P-analysis should come after the explicit formula.

\begin{align*}
\spacenewfn(n) = P(2^{\lceil \log_2 n \rceil}) = O(n)
\end{align*}

### 2017-10-24 Draft2/BigPNotation.tex
Remove all the \varnitems stuff

In order to highlight the benefit of growth arrays, I will define a new notation for time complexity. The reason for this is, for certain operations, growth arrays are only better than dynamic arrays by a constant factor. For example, dynamic arrays might allocate roughly $2\varnitems$ memory for appending $\varnitems$ items, while growth arrays would allocate roughly $\varnitems$ memory. Even though growth arrays are clearly better in this regard, the big-O space complexity for both data structures would be the same, $O(\varnitems)$.

My goal is to be able to compare the coefficients of the highest-order terms in both expressions. For example, I would like to take the ratio $\frac{2\varnitems}{\varnitems}$, see that it is $2$, and conclude that dynamic arrays allocate roughly twice as much as growth arrays for large $\varnitems$. However, big-O notation does not support this.

@@

Notice that while $O(2\varnitems) = O(\varnitems)$, $P(2\varnitems) \neq P(\varnitems)$. Thus, $P$ makes it possible to distinguish between a function that uses $\varnitems$ space and one that uses $2\varnitems$ space. \textbf{Note:} A consequence of this is that bases for logarithms must be specified in big-P notation.

>>>>

In order to highlight the benefit of growth arrays, I will define a new notation for time complexity. The reason for this is, for certain operations, growth arrays are only better than dynamic arrays by a constant factor. For example, dynamic arrays might allocate roughly $2n$ memory for appending $n$ items, while growth arrays would allocate roughly $n$ memory. Even though growth arrays are clearly better in this regard, the big-O space complexity for both data structures would be the same, $O(n)$.

My goal is to be able to compare the coefficients of the highest-order terms in both expressions. For example, I would like to take the ratio $\frac{2n}{n}$, see that it is $2$, and conclude that dynamic arrays allocate roughly twice as much as growth arrays for large $n$. However, big-O notation does not support this.

@@

Notice that while $O(2n) = O(n)$, $P(2n) \neq P(n)$. Thus, $P$ makes it possible to distinguish between a function that uses $n$ space and one that uses $2n$ space. \textbf{Note:} A consequence of this is that bases for logarithms must be specified in big-P notation.
